Robot Team III: Scotty Strikes Back

### 2024-04-30 paper/paper.tex
added update equation for  beta
updated intro some more...

1. Introductory paragraph

Inland water bodies pose a unique challenge to characterization via remote sensing imagery due to their complex spectral characteristics and small-scale spatial variability (cite reviews here). Hyperspectral imaging has emerged as a pivotal technology in this space, where increased spectral resolution enables the inversion of optically-active water quality parameters such as colored dissolved organic matter (CDOM), chlorophyll-a, blue-green algae, turbidity, and temperature from captured reflectance spectra \cite{koponen2002lake,ritchie2003remote, bonansea2015using}. Machine learning has proven to be a powerful tool in this domain, with tree-based models, support vector machines, deep neural networks, and others architectures providing data-driven approaches to map observed reflectance spectra to parameters of interest \cite{thenkabail2018hyperspectral, ghatkar2019classification,sagan2020monitoring}. However, the training and evaluation of these models demands significant volumes of coincident, in situ data. Collecting these datasets is time consuming and expensive, often requiring observations spanning decades \cite{aurin2018remote, ross2019aquasat}.

2. HSI equipped drones + in situ reference sensing --> still require that we know what to look for.

@@

3. Unsupervised methods in remote sensing and HSI
- Describe classical ML methods such as k-means, fuzzy c-means, DBSCAN, etc in remote sensing and HSI. 
    - Key limitation: how do you determine the correct number of classes?
    - This can be very hard to do without a Bayesian interpretation for the model. e.g. you have to rely on heuristics like the Silhouette diagram.
- Describe related problem of endmember extraction and spectral unmixing. Specifically there are geometric and statistical approaches to solving the problem with ML methods a popular choice for the identification of the relvant family of spectral endmembers.
- To date, much of the focus on drone-based HSI platforms has been on supervised regression of parameters OR the determination of spectral indices.
- An ideal unsupervised approach should accomplish three things
    1. Segmentation of captured imagery into a set of distinct classes
    2. Mapping of distinct spatial regions in collected imagery using the identified classes
    3. Simultaneous identification of relevant spectral endmembers

4. Generative topographic mapping 
Describe the self organizing map (SOM) as an interesting approach due to the ability to 
    - classify data 
    - maintain topographic relationship between classes which can be used to produces a lower dimensional representation of data distribution (i.e. cluster identification)
    - the weight vector for each class can be interpreted as an endmember in the original data space such that common classes can be interpreted as a family of spectral endmembers.
Cite a few uses ofthe SOM for remote sensing and HSI analysis...
However, the original SOM training algorithm relies on heuristics for controlling the updates during training and offers no clear probabilitistic interpreation. The GTM was developed to specifically address this, providing a fully Bayesian alternative to the SOM.

5. This study

@@

>>>>

%   1. Introductory paragraph

Inland water bodies pose a unique challenge to characterization via remote sensing imagery due to their complex spectral characteristics and small-scale spatial variability \cite{koponen2002lake,ritchie2003remote}. Hyperspectral imaging has emerged as a pivotal technology in this space, where the heightened spectral resolution enables the inversion of optically-active water quality parameters such as colored dissolved organic matter (CDOM), chlorophyll-a, blue-green algae, turbidity, and temperature from captured reflectance spectra \cite{brando2003satellite,moses2012estimation}.

Machine learning has proven to be a powerful tool in this domain, with tree-based models, support vector machines, deep neural networks, and others architectures providing data-driven approaches to map observed reflectance spectra to parameters of interest \cite{thenkabail2018hyperspectral,keller2018hyperspectral, ghatkar2019classification,sagan2020monitoring}. However, the calibration and evaluation of these models demands a significant volume of coincident, in situ data. Collecting paired datasets is time consuming and expensive, often relying on rare satellite overpasses spanning decades \cite{aurin2018remote, ross2019aquasat}. This requirement impedes the real-time assessment of water quality change, for example, during oil spills \cite{fingas2014review,fingas2017review}.

% 2. HSI equipped drones + in situ reference sensing --> still require that we know what to look for.

@@

% 3. Unsupervised methods in remote sensing and HSI
- Describe classical ML methods such as k-means, fuzzy c-means, DBSCAN, etc in remote sensing and HSI. 
    - Key limitation: how do you determine the correct number of classes?
    - This can be very hard to do without a Bayesian interpretation for the model. e.g. you have to rely on heuristics like the Silhouette diagram.
- Describe related problem of endmember extraction and spectral unmixing. Specifically there are geometric and statistical approaches to solving the problem with ML methods a popular choice for the identification of the relvant family of spectral endmembers.
- To date, much of the focus on drone-based HSI platforms has been on supervised regression of parameters OR the determination of spectral indices.
- An ideal unsupervised approach should accomplish three things
    1. Segmentation of captured imagery into a set of distinct classes
    2. Mapping of distinct spatial regions in collected imagery using the identified classes
    3. Simultaneous identification of relevant spectral endmembers

% 4. Generative topographic mapping 
Describe the self organizing map (SOM) as an interesting approach due to the ability to 
    - classify data 
    - maintain topographic relationship between classes which can be used to produces a lower dimensional representation of data distribution (i.e. cluster identification)
    - the weight vector for each class can be interpreted as an endmember in the original data space such that common classes can be interpreted as a family of spectral endmembers.
Cite a few uses ofthe SOM for remote sensing and HSI analysis...
However, the original SOM training algorithm relies on heuristics for controlling the updates during training and offers no clear probabilitistic interpreation. The GTM was developed to specifically address this, providing a fully Bayesian alternative to the SOM.

% 5. This study

@@

\begin{equation}
    \frac{1}{\beta_{\text{new}}} = \frac{1}{ND} \sum\limits_{n}^{N} \sum\limits_{k}^{K} R_{kn} \left\lVert \psi_k - \mathbf{x}_n \right\rVert^2
\end{equation}
